### 2015-03-16 Kokkos_PG_HierarchicalParallelism.tex
Added intro to Hierarchical Parallelism chapter.

\chapter{Hierarchical Parallelism}\label{C:Hierarchical}
Node architectures on modern high-performance computers are characterized by ever more \emph{hierarchical parallelism}. 
A level in the hierachy is determined by the hardware resources which are shared between compute units at that level. 
Higher levels in the hierarchy also have access to all resources in its branch at lower levels of the hierarchy.
This concept is orthogonal to the concept of heterogeneity. 
For example, a node in a typical CPU-based cluster consists of a number of multicore CPUs.  Each core supports one or more hyperthreads, and each hyperthread can execute vector instructions.
This means there are 4 levels in the hierarchy of parallelism: 
\begin{enumerate}
\item CPU sockets share access to the same memory and network resources,
\item cores within a socket typically have a shared last level cache (LLC), 
\item hyperthreads on the same core have access to a shared L1 (and L2) cache and they submit instructions to the same execution units, and
\item vector units execute a shared instruction on multiple data items.
\end{enumerate}
GPU-based systems also have a hierarchy of 4 levels:
\begin{enumerate}
\item Multiple GPUs in the same node share access to the same host memory and network resources, 
\item core clusters (e.g. the SMs on an NVIDIA GPU) have a shared cache and access to the same high bandwidth memory on a single GPU, 
\item threads running on the same core cluster have access to the same L1 cache and scratch memory and they are 
\item grouped in so called Warps or Wave Fronts within which threads are always synchronous and can collaborate more closely for example via direct register swapping. 
\end{enumerate}
Kokkos provides a number of abstract levels of parallelism which at compile time or run time are mapped to the appropriate hardware features. 
This mapping is not necessarily static or predefined; it may differ for each kernel.
Furthermore, some mapping decisions happen at run time. 
This enables adaptive kernels which map work to different hardware resources depending on the work set size.
While Kokkos provides defaults and suggestions, the optimal mapping can be algorithm dependent. 
Hierarchical parallelism is accessible through execution policies.

\section{Thread Teams Basics}

@@

\subsection{Creating a Policy Instance}

@@

\subsection{Basic Kernels}
Within the kernel the team policy member type provides the necessary functionality to use teams. 
It allows access to thread identifiers such as league rank and size and team rank and size, as well as team synchronous actions such as team barriers, reductions and scans.
\begin{lstlisting}
using Kokkos::TeamPolicy;
using Kokkos::parallel_for;

@@

\section{Team Scratch Pads}\label{S:Hierarchical:Scratch}

@@

\section{Nested Parallelism}

@@

\subsection{Team and Vector Loops}

>>>>

\chapter{Hierarchical Parallelism}\label{C:Hierarchical}

This chapter explains how to use Kokkos to exploit multiple levels of shared-memory parallelism.
These levels include thread teams, threads within a team, and vector lanes.
You may nest these levels of parallelism,
and execute \lstinline!parallel_for!, \lstinline!parallel_reduce!, or \lstinline!parallel_scan! at each level.
The syntax differs only by the execution policy,
which is the first argument to the \verb!parallel_*! operation.
Kokkos also exposes a ``scratch pad'' memory which all threads within a team may share.

\section{Motivation}\label{S:Hierarchical:Motivation}

Node architectures on modern high-performance computers are characterized by ever more \emph{hierarchical parallelism}. 
A level in the hierachy is determined by the hardware resources which are shared between compute units at that level. 
Higher levels in the hierarchy also have access to all resources in its branch at lower levels of the hierarchy.
This concept is orthogonal to the concept of heterogeneity. 
For example, a node in a typical CPU-based cluster consists of a number of multicore CPUs.  Each core supports one or more hyperthreads, and each hyperthread can execute vector instructions.
This means there are 4 levels in the hierarchy of parallelism: 
\begin{enumerate}
\item CPU sockets share access to the same memory and network resources,
\item cores within a socket typically have a shared last level cache (LLC), 
\item hyperthreads on the same core have access to a shared L1 (and L2) cache and they submit instructions to the same execution units, and
\item vector units execute a shared instruction on multiple data items.
\end{enumerate}
GPU-based systems also have a hierarchy of 4 levels:
\begin{enumerate}
\item Multiple GPUs in the same node share access to the same host memory and network resources, 
\item core clusters (e.g. the SMs on an NVIDIA GPU) have a shared cache and access to the same high bandwidth memory on a single GPU, 
\item threads running on the same core cluster have access to the same L1 cache and scratch memory and they are 
\item grouped in so called Warps or Wave Fronts within which threads are always synchronous and can collaborate more closely for example via direct register swapping. 
\end{enumerate}
Kokkos provides a number of abstract levels of parallelism,
which it maps to the appropriate hardware features.
This mapping is not necessarily static or predefined; it may differ for each kernel.
Furthermore, some mapping decisions happen at run time. 
This enables adaptive kernels which map work to different hardware resources depending on the work set size.
While Kokkos provides defaults and suggestions, the optimal mapping can be algorithm dependent. 
Hierarchical parallelism is accessible through execution policies.

\section{Thread teams}\label{S:Hierarchical:Teams}

@@

\subsection{Creating a Policy Instance}\label{SS:Hierarchical:Teams:Policy}

@@

\subsection{Basic Kernels}\label{SS:Hierarchical:Teams:Kernels}

Within the kernel the team policy member type provides the necessary functionality to use teams. 
It allows access to thread identifiers such as league rank and size and team rank and size, as well as team synchronous actions such as team barriers, reductions and scans.
\begin{lstlisting}
using Kokkos::TeamPolicy;
using Kokkos::parallel_for;

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Team scratch pad memory}\label{S:Hierarchical:Scratch}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Nested parallelism}\label{S:Hierarchical:Nested}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Team and vector loops}\label{SS:Hierarchical:Nested:Loops}
